---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# forstringr

<!-- badges: start -->
<!-- badges: end -->

The goal of `forstringr` is to enable complex string manipulation in R possible. The package combines the power of `stringr`, and other manipulation packages such as `dplyr` and `tidyr`. Just like in the `stringr` package, every function in `forstringr` begins with `str_`.

## Installation

You can install the development version of `forstringr` from [GitHub](https://github.com/) with

``` r
devtools::install_github("gbganalyst/forstringr")
```


## `str_left()`

Given a character vector, `str_left()` returns the left side of a string. For examples:

```{r example1}
library(forstringr)

str_left("Nigeria")

str_left("Nigeria", n = 3)

str_left(c("Female", "Male", "Male", "Female"))
```

## `str_right()`

Given a character vector, `str_right()` returns the right side of a string. For examples:

```{r example2}

str_right("July 20, 2022", 4)

str_right("Sale Price", n = 5)
```

## `str_mid()`

Just like in the Microsoft Excel, the `str_mid()`returns a specific number of characters from a text string, starting at the position you specify, based on the number of characters you specify.

```{r}
str_mid("Super Eagle", 7, 5)

str_mid("Oyo Ibadan", 5, 6)
```

## `str_split_extract()`

If you want to split up a string into pieces and extract the results using a specific index position, then, you will use `str_split_extract()`. You can interpret it as interpret it as follows:

Given a character string, `S`, extract the element at a given position, `k`, from the result of splitting `S` by a given pattern, `m`. For example:

```{r}
top_10_richest_nig <- c("Aliko Dangote", "Mike Adenuga ", "Femi Otedola", "Arthur Eze", "Abdulsamad Rabiu", "Cletus Ibeto", "Orji Uzor Kalu", "ABC Orjiakor", "Jimoh Ibrahim", "Tony Elumelu")

first_name <- str_split_extract(top_10_richest_nig, " ", 1)

first_name
```



